---
title: "R Basics"
author: "Matthias Uckert"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_notebook:
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath(here::here())) 
knitr::opts_chunk$set(root.dir = normalizePath(here::here()))
```

```{r include=FALSE}
library(plotly)
library(tidyverse)
library(htmlwidgets)
library(openxlsx)
library(janitor)
library(furrr)
library(scales)
library(ggthemes)
library(patchwork)
options(scipen = 999)
source("1_code/functions/f-02_business_case.R", encoding = "UTF-8")
```

# Define Input Functions

## Var. Cost Curve

First we define a Function that relates our variable cost to the our
sales volume. Specifically, we assume that for 20% additional units
(base: 500) that we source, we get a 5% discount from our supplier, up
to a maximum discount of 25%

```{r}
.vc <- 100
.vol <- 6000
.vol_base <- 1500
.vol_cap <- 5000
.vol_dis <- 500
.per_dis <- .5
.disc <- .05
fnc_var_cost <- function(.vc, .vol, .vol_base, .vol_cap, .vol_dis = NULL, .per_dis = NULL, .disc) {
  vol_disc_ <- ifelse(!is.null(.vol_dis), .vol_dis, .vol_base * .per_dis)
  vol_use_ <- dplyr::case_when(
    .vol < .vol_base ~ .vol_base,
    .vol > .vol_cap ~ .vol_cap,
    TRUE ~ .vol
  )
  disc_ <- floor((vol_use_ - .vol_base) / vol_disc_) * .disc
  .vc * (1 - disc_) * .vol
}
fnc_var_cost(2250, 550)
```

We define the variable cost as 1 to show the degression effects in
percent

```{r echo=FALSE, message=FALSE, warning=FALSE}
tab_vc <- tibble(
  vc = 1, 
  vol = seq(1000, 10000, 5), 
  vol_base = 1500,
  vol_cap = 7500, 
  vol_dis = 750, 
  per_dis = NULL, 
  disc = .05
  ) %>%
  mutate(
    var_cost = fnc_var_cost(vc, vol, vol_base, vol_cap, vol_dis, per_dis, disc),
    unit_vc = var_cost / vol
    ) 

gg_vc_unit <- plot_points(tab_vc, vol, unit_vc, "Volume", "Unit Variable Cost (Volume)")
gg_vc_tot  <- plot_points(tab_vc, vol, var_cost, "Volume", "Total Variable Cost (Volume)")

gg_vc_unit + gg_vc_tot


```

## Fix Cost Curve

Next we define a step Function, that doubles our directly attributable
fixed for every additional 750 units we produce

```{r message=FALSE, warning=FALSE}
.fc <- 50000
.vol <- 6000
.vol_mult <- 1000
fnc_fix_cost <- function(.fc, .vol, .vol_mult) {
  .fc * ceiling(.vol / .vol_mult)
}

fnc_fix_cost(50000, 6000, 1000)
```

We define the fix cost as 750, so the step-function will approach 1 as
the volume goes to infinity.

```{r echo=FALSE, message=FALSE, warning=FALSE}
tab_fc <- tibble(
  fc = 1, 
  vol = seq(250, 10000, 5), 
  vol_mult = 1000,
  ) %>%
  mutate(
    fix_cost = fnc_fix_cost(fc, vol, vol_mult),
    unit_fc = fix_cost / vol
    ) 

gg_fc_unit <- plot_points(tab_fc, vol, unit_fc, "Volume", "Unit Fix Cost (Volume)")
gg_fc_tot  <- plot_points(tab_fc, vol, fix_cost, "Volume", "Total Fix Cost (Volume)")

gg_fc_unit + gg_fc_tot
```



## Sales-Price Curve

Next, we define a a sigmoid function to proxy the sales price we can
achieve as a function of sales volume. We cap this function on a minimum
of 25% of our initially set price.

```{r message=FALSE, warning=FALSE}
.vol <- 1
fnc_revenue <- function(.price, .vol) {
  sigmoid <- function(x) {
    y = 1 - (1 / (1 + exp(-((x / 1000) - 4.25))))
    return(y)
  }
  .vol * .price * (sigmoid(.vol) * .4 + .6)
}
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
tab_rv <- tibble(vol = seq(1, 10000, 5)) %>%
  mutate(
    price = 1,
    rev = fnc_revenue(price, vol),
    unit_rev = rev / vol
    )

fnc_revenue(750, 1000) / 1000

gg_rev_unit <- plot_points(tab_rv, vol, unit_rev, "Volume", "Price (Volume)")
gg_rev_tot  <- plot_points(tab_rv, vol, rev, "Volume", "Revenue (Volume)")

gg_rev_unit + gg_rev_tot
```

## CM I, CM II, and Profit
```{r message=FALSE, warning=FALSE}
tab_sens <- get_sensitivity(
  .vol = 10000, .price = 750, .vc = 500, .vol_base = 2500, .vol_cap = 5000,
  .vol_dis = 500, .per_dis = NULL, .disc = .05, .fc = 25000, .vol_mult = 1000,
  .cost = 250000, .ret = "Profit"
) %>%
  dplyr::filter(input %in% c("price", "volume", "fix_cost", "var_cost", "firm_cost", "disc")) %>%
  plot_sensitivity()
```


```{r}


map_dfr(c("vol", "price", "vc", "fc"), ~ get_sensitivity(
  .vol = 10000,
  .price = 750,
  .vc = 500,
  .vol_base = 2500,
  .vol_cap = 5000,
  .vol_dis = 500,
  .per_dis = NULL,
  .disc = .05,
  .fc = 25000,
  .vol_mult = 1000,
  .cost = 250000,
  .ret = "Profit",
  .col = .x
))

get_sensitivity(
  .vol = 10000,
  .price = 750,
  .vc = 500,
  .vol_base = 2500,
  .vol_cap = 5000,
  .vol_dis = 500,
  .per_dis = NULL,
  .disc = .05,
  .fc = 25000,
  .vol_mult = 1000,
  .cost = 250000,
  .ret = "Profit",
  .col = disc
) %>% plot_profits(seq, val)


```


```{r}
generate_table <- function(.vol = NULL, .price = NULL, .vc = NULL, .vol_base = NULL, 
                           .vol_cap = NULL, .vol_dis = NULL, .per_dis = NULL,
                           .disc = NULL, .fc = NULL, .vol_mult = NULL, .cost = NULL) {
  tibble::tibble(
    vol = .vol,
    price = .price,
    vc = .vc,
    vol_base = .vol_base,
    vol_cap = .vol_cap,
    vol_dis = .vol_dis,
    per_dis = .per_dis,
    disc = .disc,
    fc = .fc,
    vol_mult = .vol_mult,
    cost = .cost
  ) %>%
    dplyr::mutate(
      var_cost = fnc_var_cost(vc, vol, vol_base, vol_cap, vol_dis, per_dis, disc),
      unit_vc = var_cost / vol,
      fix_cost = fnc_fix_cost(fc, vol, vol_mult),
      unit_fc = fix_cost / vol,
      rev = fnc_revenue(price, vol),
      unit_rev = rev / vol
    ) %>%
    dplyr::mutate(
      cm1 = rev - var_cost,
      cm2 = rev - var_cost - fix_cost,
      profit = cm2 - cost
    )
}
```


```{r}
tab_all <- generate_table(
  .vol = 10000, .price = seq(500, 1250, 5), .vc = 500, .vol_base = 2500, .vol_cap = 5000, 
  .vol_dis = 500, .per_dis = NULL, .disc = .05, .fc = 25000, .vol_mult = 1000, .cost = 250000)

gg_cm1 <- plot_profits(tab_all, price, cm1, "", "", "CM I")
gg_cm2 <- plot_profits(tab_all, price, cm2, "", "", "CM II")
gg_prf <- plot_profits(tab_all, vc, profit, "", "", "Profit")

gg_cm1
gg_cm2
gg_prf
```

```{r}
tab_all <- generate_table(
  .vol = 10000, .price = 750, .vc = 500, .vol_base = 2500, .vol_cap = 5000, 
  .vol_dis = 500, .per_dis = NULL, .disc = .05, .fc = 25000, .vol_mult = 1000, .cost = 250000
  ) %>%
  expand_grid(seq = seq(0, 2, .01))
```


# Contribution Margin & Profit

We define a Function that calculates the CM I and CM II per product

```{r}
fnc_cm <- function(.volume, .price, .var_cost, .fix_cost, .type = c("CM I", "CM II")) {
  revenue_ <- fnc_revenue(.price, .volume)
  
  if (.type == "CM I") {
    revenue_ - fnc_var_cost(.var_cost, .volume)
  } else {
    revenue_ - fnc_var_cost(.var_cost, .volume) - fnc_fix_cost(.fix_cost, .volume)
  }
}
```

## CM as a Function of Volume

Since all of our Input Factors are depended on Volume, we plot a CM I
and CM II volume-curve Let's look what happens with out base
assumptions:

-   **Price**: 899 EUR

-   **Variable Cost**: 550 EUR

-   **Fix Cost**: 80,000 EUR

```{r echo=FALSE, message=FALSE, warning=FALSE}
tab_cm <- tibble(volume = 100:10000) %>%
  mutate(
    cm1 = fnc_cm(volume, 899, 575, 20000, "CM I"),
    cm2 = fnc_cm(volume, 899, 575, 20000, "CM II"),
    unit_cm1 = cm1 / volume,
    unit_cm2 = cm2 / volume
  ) %>%
  pivot_longer(c(cm1, cm2, unit_cm1, unit_cm2)) 
```

### Unit CM I
```{r echo=FALSE, message=FALSE, warning=FALSE}
gg_unit_cm1 <- tab_cm %>%
  filter(name %in% c("unit_cm1")) %>%
  mutate(col = if_else(value < 0, "red", "green")) %>%
  plot_cm()
plotly::ggplotly(gg_unit_cm1, width = 750)
```

### Unit CM I
```{r echo=FALSE, message=FALSE, warning=FALSE}
gg_unit_cm2 <- tab_cm %>%
  filter(name %in% c("unit_cm2")) %>%
  mutate(col = if_else(value < 0, "red", "green")) %>%
  plot_cm()
plotly::ggplotly(gg_unit_cm2, width = 750)
```
### Total CM II

```{r echo=FALSE, message=FALSE, warning=FALSE}
gg_total_cm2 <- tab_cm %>%
  filter(name %in% c("cm2")) %>%
  mutate(col = if_else(value < 0, "red", "green")) %>%
  plot_cm()
plotly::ggplotly(gg_total_cm2, width = 750)
```

# Sensitivity
Next, we analyse how sensitive the contribution margin is to our 4 different input factors, 
we use the following base case scenario. 

-   **Price**: 899 EUR

-   **Variable Cost**: 550 EUR

-   **Fix Cost**: 80,000 EUR

-   **Volume**: 1,000 Units


```{r}
s_ <- seq(.01, 2.99, .01)
b_ <- fnc_cm(2250, 899, 550, 80000, "CM II")
tab_sensitivity <- tibble(base = s_) %>% 
  mutate(
    volume = fnc_cm(2250 * s_, 899, 550, 80000, "CM II") / b_,
    price = fnc_cm(2250, 899 * s_, 550, 80000, "CM II") / b_,
    var_cost = fnc_cm(2250, 899, 550 * s_, 80000, "CM II") / b_,
    fix_cost = fnc_cm(2250, 899, 550, 80000 * s_, "CM II") / b_
    ) %>%
  pivot_longer(volume:fix_cost)

tab_sensitivity %>%
  mutate(base = base - 1) %>%
  ggplot(aes(base, value)) + 
  geom_point(color = "blue", fill = "white", size = .75) + 
  facet_wrap(~ name, scales = "free_y") + 
  theme_tufte() + 
  geom_rangeframe() + 
  scale_x_continuous(labels = percent) + 
  scale_y_continuous(labels = percent) + 
  geom_hline(yintercept = 1) + 
  geom_vline(xintercept = 0) + 
  geom_vline(xintercept = 1)
  
```



# Monte Carlo Analysis
```{r message=FALSE, warning=FALSE}
tab_montecarlo <- tibble(
  price = runif(1000000, 600, 1250),
  volume = runif(1000000, 2000, 10000),
  var_cost = runif(1000000, 500, 750),
  fix_cost = runif(1000000, 60000, 100000),
) %>%
  mutate(
    revenue = fnc_revenue(price, volume),
    cm1 = fnc_cm(volume, price, var_cost, fix_cost, "CM I"),
    cm2 = fnc_cm(volume, price, var_cost, fix_cost, "CM II")
  ) %>%
  mutate(margin = cm2 / revenue)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
tab_montecarlo %>%
  mutate(col = if_else(margin < 0, "red", "green")) %>%
  ggplot(aes(margin, fill = col, alpha = .5)) +
  geom_histogram(binwidth = .001) + 
  ggplot2::scale_fill_manual(values = c("green", "red")) +
    ggthemes::geom_rangeframe() +
    ggthemes::theme_tufte() +
    ggplot2::scale_x_continuous(labels = scales::percent) +
    ggplot2::scale_y_continuous(labels = scales::comma) + 
    ggplot2::geom_vline(xintercept = 0, linetype = "dashed") + 
    ggplot2::theme(legend.position = "none")

tab_montecarlo %>%
  mutate(col = if_else(margin < 0, "red", "green")) %>%
  ggplot(aes(cm2, fill = col, alpha = .5)) +
  geom_histogram(bins = 1000) + 
  ggplot2::scale_fill_manual(values = c("green", "red")) +
    ggthemes::geom_rangeframe() +
    ggthemes::theme_tufte() +
    ggplot2::scale_x_continuous(labels = scales::comma) +
    ggplot2::scale_y_continuous(labels = scales::comma) + 
    ggplot2::geom_vline(xintercept = 0, linetype = "dashed") + 
    ggplot2::theme(legend.position = "none")
```

```{r}
tibble(
  name = c("CM II < 0", "CM II > 0", "Mean"),
  value = c(
    percent(nrow(filter(tab_montecarlo, cm2 < 0)) / nrow(tab_montecarlo), .01),
    percent(nrow(filter(tab_montecarlo, cm2 > 0)) / nrow(tab_montecarlo), .01),
    comma(mean(tab_montecarlo$cm2))
  )
) %>% bind_rows(
  enframe(comma(quantile(tab_montecarlo$cm2, probs = c(seq(.25, .75, .25)))))
)




```



```{r}
tab_montecarlo %>%
  filter(between(margin, quantile(margin, probs = .50) - .0001, quantile(margin, probs = .50) + .0001)) %>%
  DT::datatable() %>%
  DT::formatCurrency(
    columns = c("price", "volume", "var_cost", "fix_cost", "revenue", "cm1", "cm2"), 
    currency = "", 
    interval = 3, 
    mark = ","
    ) %>%
  DT::formatPercentage("margin", digits = 2)
```

```{r}
tab_montecarlo %>%
  filter(between(margin, quantile(margin, probs = .25) - .0001, quantile(margin, probs = .25) + .0001)) %>%
  DT::datatable() %>%
  DT::formatCurrency(
    columns = c("price", "volume", "var_cost", "fix_cost", "revenue", "cm1", "cm2"), 
    currency = "", 
    interval = 3, 
    mark = ","
    ) %>%
  DT::formatPercentage("margin", digits = 2)
```

```{r}
tab_montecarlo %>%
  filter(between(margin, quantile(margin, probs = .75) - .0001, quantile(margin, probs = .75) + .0001)) %>%
  DT::datatable() %>%
  DT::formatCurrency(
    columns = c("price", "volume", "var_cost", "fix_cost", "revenue", "cm1", "cm2"), 
    currency = "", 
    interval = 3, 
    mark = ","
    ) %>%
  DT::formatPercentage("margin", digits = 2)
```
